---
title: "`r params$lhd`"
author: "Ryan Thornburg, Carolina Data Desk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    css: styles.css
params:
  county: default
  lhd: default   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(scales))


lhd <- params$lhd


pagedata <- wide_CompiledLHDExpenditures %>% 
  filter(county_name == county) 

# Is it single county or multi-county?
what<- case_when(
    pagedata$lhd_area_type=="single" ~ "County is",
    pagedata$lhd_area_type=="multi-county" ~ "counties are",
    TRUE ~ ""
  )
  
  
  



getverb <- function(mychange){
  myverb <- case_when(
    mychange > 0 ~ "rose",
    mychange < 0 ~ "fell",
    TRUE ~ "was unchanged"
  )
  return(myverb)
}


inflspendchange_raw <- pagedata$expenditures_infl_2018 - pagedata$expenditures_infl_2010
inflspendchange_pct <- (pagedata$expenditures_infl_2018 - pagedata$expenditures_infl_2010) / pagedata$expenditures_infl_2010

popchange_pct <- (pagedata$population_2018 - pagedata$population_2010) / pagedata$population_2010

expenditures_infl_per_capita_change_pct <- (pagedata$expenditures_infl_per_capita_2018 - pagedata$expenditures_infl_per_capita_2010) / pagedata$expenditures_infl_per_capita_2010





```


`r county` `r what` served by the `r lhd`. From 2010 to 2018 the department's spending `r getverb(inflspendchange_raw)` by `r scales::percent(abs(inflspendchange_pct))`, when adjusted for inflation. During that same time, the population served by the `r lhd` `r getverb(popchange_pct)` by `r scales::percent(abs(popchange_pct))`. When taking that population change into account, the per-person spending `r getverb(expenditures_infl_per_capita_change_pct)` `r scales::percent(abs(expenditures_infl_per_capita_change_pct))`.    


In 2010, the `r lhd` ranked `r pagedata$per_cap_rank_2010` in per-person spending among the `r denominator` health departments for which we have complete data. In 2018, it ranked `r pagedata$per_cap_rank_2018`.

Across the state, county-level spending on public health dropped 22% from 2010-2018 when adjusted for inflation and the change in population.


```{r spending_graph, warning=FALSE, message=FALSE}

page_graph_data <- pagedata %>%
  select(lhd_name, starts_with("expenditures_infl_per_capita_")) %>%
  pivot_longer(cols = c(starts_with("expenditures_infl_per_capita_")),
               names_to = "year",
               names_prefix = "expenditures_infl_per_capita_",
               values_to = "avg_expenditures_infl_per_capita")



  ggplot() +
      geom_line(data=page_graph_data,
            aes(x=year,
                y=avg_expenditures_infl_per_capita,
                group = 1,
                col = params$lhd)
            ) +
    geom_line(data = graphmeans, 
         aes(x=year, 
             y=avg_expenditures_infl_per_capita,
             group = 1,
             col ="Statewide Average")) +

    labs(title="Local Health Department Spending",
         caption = "Source: Carolina Data Desk. https://github.com/carolinadatadesk/NC-Public-Health-Data")+
    scale_x_discrete(name="Years") + 
    scale_y_continuous(name = "Per Person", labels = dollar) +
    aptheme::theme_ap() +
    theme(text=element_text(family="Arial")) + 
    theme(legend.position = 'bottom')

```

